---
title: "TampaParks_SO_PREP"
author: "Gabriel"
date: "2024-04-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(tidyverse)
library(tidycensus)
library(dplyr)
library(ggplot2)
library(kableExtra)
install.packages('tmap', dependencies = TRUE)
library(scales)

options(scipen=999)
options(tigris_class = "sf")

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

#Load API KEY
census_api_key("3aedcfd189958d66a813a29a4b33580e78657644", overwrite = TRUE)
#Set CRS

crs='EPSG:3425'

```

* The purpose of this study is to gain insight into the relationship between the residents of Tampa, Florida and the City of Tampa's parks service. I am approaching this project from the perspective of a non-profit organization interested in the putting on events that supply food and home goods to families at a scale that would be deeply impact their yearly cost-of-living expenses. We also want to advocate for more funds to be programmed at the top 3 parks that serve our targe demographic. Thanks to detailed from the American Survey Data 5-Year Estimates for 2022, it is possible to compile the counts of varying household and householder characteristics at a census block group level. We will test for different target audiences and compare visualizations of the outputs.
  The project is based on data that is publicly available. I will be using a parks polygon shapefile from the City of Tampa Geohub website: https://city-tampa.opendata.arcgis.com/ . I will turn them into points that are geopositioned at the weighted center of the polygon (centroids). The relationship between parks and people is partly defined by access. In this project, I define accessibility as being able to walk 15 minutes to the park. Tampa is not a public transportation oriented city, and due to the heat and humidity a walk longer tha
  The investigation of this project is mathematically possible through the application of the location-allocation problem from the field of spatial optimization. In simple terms, we start with series of 'demand' and 'facility' nodes. We will then decide on a distance that will be used to create a coverage matrix. A coverage matrix uses the binary values of 1 and 0 to symbolize whether a facility with a radius of a given distance covers a demand node. The coverage matrix is used to generate the formula that solves our optimization problem, which will also take into account how I value the demand node and my budget that determines the desired count of facilities. The value of our demand nodes will be based on the sum of specific census variables. The desired count of parks that our group is interested in investing in improving is five. 
  Below, we begin loading the data, exploring it, defining variables, and preparing the final centroid shapefiles to bring into ArcPro for the final analysis. We are looking to sum the count of single parents as a demand value, as well as the count of families with incomes below 30K. The Federal HUD and State Income Limit in 2024 for a 4-person household at 30 percent of the AMI is $31,200. Lastly, we will look at educational attainment of the population aged 25 and above. 

```{r}

#Load Variables
acs_variable_list.22 <- load_variables(2022, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)


acs22tr_vars <-c(
               "B25002_003E", #TotalVacant
               "B21004_001E", #Median income in the past 12 months (in 2022 inflation-adjusted dollars
               "B09021_023E", #65 years and over Lives alone
               "B25136_001E", #Total Housing Units
               "B25071_001E", #Median gross rent as a percentage of household income
               "B05006_143E", #Total:!!Americas:!!Latin America:!!Caribbean:!!Cuba
               "B19057_002E", #Total:!!With public assistance income
               "B25035_001E", #Median year structure built
               "B25077_001E", #med values
               "B05010_002E", #Ratio of Income to Poverty Level under 1.00
               "B02001_003E", #Race Black 
               "B02001_008E", #Race - 2 or more
               "B03002_012E", #Total Hispanic
               "B17010_011E", #bg Male householder, no spouse present:!!With related children of the householder under 18 years
               "B17010_017E", #bg Female householder, no spouse present:!!With related children of the householder under 18 years:
               "B17017_003E", # bg Estimate!!Total:!!Income in the past 12 months below poverty level:!!Family household
               "B19101_002E", #bg family income Estimate!!Total:!!Less than $10,000
               "B19101_003E", #bg family income $10,000 to $14,999
               "B19101_004E", #bg family income Estimate!!Total:!!$15,000 to $19,999
               "B19101_005E", #bg  family income !$20,000 to $24,999
               "B19101_006E", #bg family income Total:!!$25,000 to $29,999
               "B15003_023E", #bg educational attainment (EA) 25+ MAsters
               "B15003_024E", #bg EA 25+ professinal degrees
               "B15003_025E"  #bg EA25+ doctorates
)

acs22bg_vars<-c("B17010_011E", #bg Male householder, no spouse present:!!With related children of the householder under 18 years
               "B17010_017E", #bg Female householder, no spouse present:!!With related children of the householder under 18 years:
               "B17017_003E", # bg Estimate!!Total:!!Income in the past 12 months below poverty level:!!Family household
               "B19101_002E", #bg family income Estimate!!Total:!!Less than $10,000
               "B19101_003E", #bg family income $10,000 to $14,999
               "B19101_004E", #bg family income Estimate!!Total:!!$15,000 to $19,999
               "B19101_005E", #bg  family income !$20,000 to $24,999
               "B19101_006E", #bg family income Total:!!$25,000 to $29,999
               "B15003_023E", #bg educational attainment (EA) 25+ MAsters
               "B15003_024E", #bg EA 25+ professinal degrees
               "B15003_025E", #bg EA25+ doctorates)
               "B25002_003E", #TotalVacant
               "B09021_023E", #65 years and over Lives alone
               "B25136_001E", #Total Housing Units
               "B25071_001E", #Median gross rent as a percentage of household income
               "B19057_002E", #Total:!!With public assistance income
               "B25035_001E", #Median year structure built
               "B25077_001E", #med values
               "B02001_003E", #Race Black
               "B02001_008E", #Race - 2 or more
               "B03002_012E" #Total Hispanic
               
               
               
)

hillsborough_tr22 <- get_acs(geography = "tract",
                         year = 2022, 
                         variables = acs22tr_vars, 
                         geometry = TRUE, 
                         state = "FL", 
                         county = "Hillsborough", 
                         output = "wide")                
            
hillsborough_tr22 <- hillsborough_tr22%>%
  rename(
         tot_vacant22 = B25002_003E,
         med_inc22 = B21004_001E,
         up65_alone22 = B09021_023E,
         tot_housingunits22 =B25136_001E,
         gross_v_inc_perc22 = B25071_001E,
         tot_fb_cuba22 = B05006_143E,
         tot_pa_inc22 = B19057_002E,
         med_yb22 = B25035_001E,
         med_values22 = B25077_001E,
         count_under_pov= B05010_002E,
         count_black=B02001_003E,
         count_hispanic=B03002_012E,
         count_mixedrace=B02001_008E
         )%>%
  mutate(vacpct22 = (tot_vacant22/tot_housingunits22), count_poc= (count_black + count_hispanic + count_mixedrace),sum_singleparents=    (B17010_011E+B17010_017E), sum_EA_MsDr=(B15003_023E +B15003_024E+ B15003_025E), sum_fam_below_30k= (B19101_002E+B19101_003E+B19101_004E+B19101_005E+B19101_006E)) %>%
  st_as_sf(crs = crs)


#testing why shapefile isnt working

hillsborough2_bg22 <- get_acs(geography = "block group",
                         year = 2022, 
                         variables = acs22bg_vars, 
                         geometry = TRUE, 
                         state = "FL", 
                         county = "Hillsborough", 
                         output = "wide")   
#convert the block groups to centroids before exporting as shapefiles

hillsborough2_bg22 <- hillsborough2_bg22%>%
  rename(tot_vacant22 = B25002_003E,
         up65_alone22 = B09021_023E,
         tot_housingunits22 =B25136_001E,
         gross_v_inc_perc22 = B25071_001E,
         
         tot_pa_inc22 = B19057_002E,
         med_yb22 = B25035_001E,
         med_values22 = B25077_001E,
         count_black=B02001_003E,
         count_hispanic=B03002_012E,
         count_mixedrace=B02001_008E
         )%>%
  mutate(vacpct22 = (tot_vacant22/tot_housingunits22), count_poc= (count_black + count_hispanic + count_mixedrace),sum_singleparents=    (B17010_011E+B17010_017E), sum_EA_MsDr=(B15003_023E +B15003_024E+ B15003_025E), sum_fam_below_30k= (B19101_002E+B19101_003E+B19101_004E+B19101_005E+B19101_006E)) %>%
  st_as_sf(crs = crs)

#Convert the polygons to points for the analysis
hillsborough2_bg22<-st_centroid(hillsborough2_bg22[!st_is_empty(hillsborough2_bg22), ])

#remove the Margins from the data table
hillsborough2_bg22 <-hillsborough2_bg22[, !grepl("M$", names(hillsborough2_bg22))]





```
 
**About Tampa Parks

In the dataset that I acquired, there are 200 public parks owned by the city of tampa. Of those 200, there are 3 cemeteries, 3 public schools, 1 beach, 1 zoo, 1 golf course, and 191 'parks'. The median size of the parks is 3.6 acres but the mean is 14.5 acres. Because the mean is influenced by large parks, we will use the median of 3.5 acres to filter our dataset. Given more time, I would compare the differences between the coverage of small parks and larger parks using the objective function of our spatial optimization model to discuss what populations are serviced by what sized parks. It could open up a discussion about equity in the City of Tampa.

Taking the summary of the ACRES variable in the parks dataset, the 3rd quartile is 8.6 acres. Because we will be using the distancefrom a centroid to define coverage, if the demand nodes are really far from the center of large park compared to a small one it will be problematic. Thus, we will further filter the dataset to include parks below the 8.6 acre value. The reason being that we do not want to operate or advocate for changes to large parks, which we will define as anything beyond the 3rd quartile of existing Tampa park sizes. The remaining 49 parks will be turned into centroids that become the facility nodes in the spatial optimization problem. 
```{r}
#load the tampa parks 

tampa_parks<-st_read("C:/Users/Gabriel/Documents/GitHub/Tampa/data/parkpolygons/Park_Polygons.shp")

tampa_parks<- tampa_parks %>%
              filter(OWNER=='City of Tampa', ACRES > 3.5,ACRES < 8.6, FEATURECOD=='Park')
tampa_parks<-st_centroid(tampa_parks[!st_is_empty(tampa_parks), ])

```

```{r}
#exportshapefile of the census block groups
## esri field names must be less than 10 characters
if(file.exists("data/hillsborough22_bg/hillsborough_bg22.shp")) {
  file.remove(list.files("data/hillsborough22_bg", full.names = TRUE))
}

if(file.exists("data/hillsborough22_bg/hillsborough2_bg22.shp")) {
  file.remove(list.files("data/hillsborough22_bg", full.names = TRUE))
}

#st_write(hillsborough_bg22, "data/hillsborough22_bg/hillsborough_bg22.shp", delete_layer = TRUE)
st_write(hillsborough2_bg22, "data/hillsborough22_bg/hillsborough2_bg22.shp", delete_layer = TRUE)

#write the Tampa Parks Shapefile
st_write(tampa_parks, "data/filtered_CoT_parks/tampa_parks.shp", delete_layer = TRUE)

```




*Conclusion

In taking a survey of what parks in Tampa serve what groups of people, the city could potentially consider more careful planning initiatives on their public land. Architects and other types of designers that want to impact the lives of marginalized people through beautification and participatory place-making techniques could also find our work meaningful. 
  
  
  